Wiggling the bandwidth

### 2023-01-28 paper.tex
add and fix parts on splitting and iterating

% notation and basics
As announced, let us fix the Laplacian kernel function
\[ k_\gamma(x,x') = \exp\left(-\frac{\| x - x' \|}{\gamma} \right) \]
with bandwidth $\gamma > 0$.
The convention here is that smaller bandwidth actually decreases the 
area under the curve $x \mapsto k_0(x) \coloneqq k(0,x)$ and vice-versa.
This whole note is based on the observation that the kernel ridgeless 
estimators use the bandwidth twice in distinct places:
once in the computation of the \emph{dual coefficients}
\[ c_i = (\mathbf K_\gamma^\dagger Y)_i \]
where the kernel matrix depends on it and then
in the evaluation part
\[ \sum_i c_i k_\gamma(x,x_i) \]
where the kernel is computed to check similarities between sample 
points $x_i$ and unseen points $x$.
Note that the dependence of the dual coefficients on the bandwidth is 
not even guaranteed to be continuous and generally hard to understand
due to the appearing pseudo-inverse.
However, the evaluation part offers a different perspective, if we 
fix the dual coefficients $c_i$ and an unseen point $x$, then
the relationship
\[ \gamma \mapsto \sum_{i=1}^n c_i k_\gamma(x,x_i) \]
is smooth: indeed, $\gamma$ appears only as a factor to the non-smooth
norm in the exponential.

@@

% computational trick
Note that the comparison of the different bandwidths can be computed 
in a highly vectorized way:
Let $\mathbf D$ denote the distance matrix of 2-norms between $\Tr_1$ and 
$\Tr_2$.
Once it is computed, we can test a new wiggled estimator 
for \emph{any} other bandwidth $\gamma$ by running
the vectorized computation
\[ \big( \mathbf c^{\gamma_0} \exp(-(1/\gamma) \mathbf D) \big)^T \]
yielding the new predictions of the inputs in $\Tr_2$.
Thus, the big hard part of the computation, i.e. finding $\mathbf D$,
needs to be done only once for size of $\Gamma$.

@@

% questions
Summing up, here is the questions we aim to provide experimental 
evidence for:
\begin{enumerate}
    \item How close are wiggled estimators and standard estimators with 
        the wiggle bandwidth? 
        What is the radius of $\gamma$'s around $\gamma_0$ such that one 
        can move knowledge from one to the other? 
        How to choose the wiggle range $\Gamma$?
    \item What would be a good point to split the training set in the 
        context of wiggle search?
        When does wiggle search provide an advantage over standard 
        estimators (or if at all)?
    \item Given that we move away from the interpolating state 
        (in a smooth and controlled manner) does wiggling compare to 
        Tikhonov?
        What happens with the norm/complexitie of a wiggled estimator
        compared to the standard one?
    \item Does an iterated wiggle search find the best bandwidth?
        Is it less expensive than classical model selection techniques?
\end{enumerate}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=\textwidth]{split}
    \caption{split}
    \label{fig:split}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment: Does it compare to Tikhonov?}

\begin{figure}[htp]
    \centering
    \includegraphics[width=\textwidth]{tikhonov}
    \caption{tikhonov}
    \label{fig:tikhonov}
\end{figure}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=\textwidth]{iterative_algo}
    \caption{iterative algo}
    \label{fig:iterative_algo}
\end{figure}

>>>>

% notation and basics
As announced, let us fix the Laplacian kernel function
\[ k_\gamma(x,x') = \exp\left(-\gamma \| x - x' \| \right) \]
with bandwidth $\gamma > 0$.
The convention here is that smaller bandwidth actually increases the 
area under the curve $x \mapsto k_0(x) \coloneqq k(0,x)$ and vice-versa.
(different conventions everywhere!).
This whole note is based on the observation that the kernel ridgeless 
estimators use the bandwidth twice in distinct places:
once in the computation of the \emph{dual coefficients}
\[ c_i = (\mathbf K_\gamma^\dagger Y)_i \]
where the kernel matrix depends on it and then
in the evaluation part
\[ \sum_i c_i k_\gamma(x,x_i) \]
where the kernel is computed to check similarities between sample 
points $x_i$ and unseen points $x$.
Note that the dependence of the dual coefficients on the bandwidth is 
not even guaranteed to be continuous and generally hard to understand
due to the appearing pseudo-inverse.
However, the evaluation part offers a different perspective, if we 
fix the dual coefficients $c_i$ and an unseen point $x$, then
the relationship
\[ \gamma \mapsto \sum_{i=1}^n c_i k_\gamma(x,x_i) \]
is smooth: indeed, $\gamma$ appears only as a factor to the non-smooth
norm in the exponential.

@@

% computational trick
Note that the comparison of the different bandwidths can be computed 
in a highly vectorized way:
Let $\mathbf D$ denote the distance matrix of 2-norms between $\Tr_1$ and 
$\Tr_2$.
Once it is computed, we can test a new wiggled estimator 
for \emph{any} other bandwidth $\gamma$ by running
the vectorized computation
\[ \big( \mathbf c^{\gamma_0} \exp(-\gamma \mathbf D) \big)^T \]
yielding the new predictions of the inputs in $\Tr_2$.
Thus, the big hard part of the computation, i.e. finding $\mathbf D$,
needs to be done only once for size of $\Gamma$.

@@

% questions
Summing up, here is the questions we aim to provide experimental 
evidence for:
\begin{enumerate}
    \item How close are wiggled estimators and standard estimators with 
        the wiggle bandwidth? 
        What is the radius of $\gamma$'s around $\gamma_0$ such that one 
        can move knowledge from one to the other? 
        How to choose the wiggle range $\Gamma$?
    \item What would be a good point to split the training set in the 
        context of wiggle search?
        When does wiggle search provide an advantage over standard 
        estimators (or if at all)?
    \item Does an iterated wiggle search find the best bandwidth?
        Is it less expensive than classical model selection techniques?
    %\item Given that we move away from the interpolating state 
    %    (in a smooth and controlled manner) does wiggling compare to 
    %    Tikhonov?
    %    What happens with the norm/complexity of a wiggled estimator
    %    compared to the standard one?
\end{enumerate}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{split}
    \caption{
        Figuring out the sweet spot for the train split on real data.
        "Baseline" always refers to the standard estimator with the 
        base bandwidth on the full train set.
        The test MSE curves refer to the best wiggled bandwidth. }
    \label{fig:split}
\end{figure}

The last experiment gave hope that wiggle search can work well.
In this part we will try to figure out the sweet spot between computational
savings and performance gain.
We move to real data now.

We choose the MNIST binary classification task with digits 4 and 9
(the hardest to distinguish) with fixed sample of size $N_{\Tr} = 1000$.
The set $\Gamma = \Gamma(\gamma_0)$ is being chosen at each base bandwidth
as a small logarithmic scale clustered at the center $\gamma_0$ 
(contains 30 bandwidths) and we fix $\gamma_0 = 1/784$ which is the 
\texttt{sklearn} default choice (based purely on dimensionality).
The splits are percentages $p$ of $N_{\Tr}$.
All plotted values are with respect to $p \in [0.3, 0.95]$:
sample error, error on held-out set, error on test set, as well as the 
chosen bandwidths and the fitting times.
See \cref{fig:split} for the results.

We summarize the observations:
\begin{itemize}
    \item The more data we are considering, the bigger we choose the 
        bandwidth parameter (e.g. the smaller the area under the kernel
        function will be) as one would expect.
    \item We achieve equivalent performance just by using only 70\% of the
        training set plus wiggling.
        After that point we see only improvements in favor of wiggling.
\end{itemize}

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{iterative_algo}
    \caption{iterative algo}
    \label{fig:iterative_algo}
\end{figure}

We have already seen in the first experiment that the direction given by
wiggle search seems to be the right one.
In this experiment we are going to test the iterated wiggle search as 
described above: fit least squares, wiggle search, refit with the best bandwidth, 
and so on.
If the chosen bandwidth converges to something, we are likely to have hit an 
optimum or at least something close to one.

The experimental setup is the same as above.
MNIST binary, initial bandwidth $\gamma_0 = 1/784$ and by our previous 
findings, train split $p = 0.85$.
We plot both the wiggled estimator's MSE with best bandwidth and that 
of refit standard estimator with exactly that MSE.
We also plot the history of the best bandwidth found.
The algorithm was run for more than 9 iterations but it converged 
and we cut the plot.
See the results in \cref{fig:iterative_algo}.

Summary:
\begin{itemize}
    \item The algorithm converges and decreases the MSE as hoped for.
    \item Remarkably, the curves for wiggled and refit estimators are
        qualitatively the same! 
        This gives hope for finding good theoretical guarantees for this
        method.
    \item The wiggled estimators reliably outperform the standard ones
        even though they are not fit on the full training set.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Experiment: Does it compare to Tikhonov?}
%
